Query Optimization for Dynamic Imputation

### 2017-07-14 paper/algorithm.tex
small fixes
paper: add table for notation

\section{Algorithm}
In order to correctly plan around missing values, we first extend the set of relational algebra operators (selection $\sigma$, projection $\pi$, join $\bowtie$, and group-by/aggregate $g$) 
with two new operators (\Cref{sec:operators}). We then define the search space (\Cref{sec:search-space}) of plans to encompass non-trivial operations, such as joins and 
aggregations. Soundness of query execution is provided by our main design invariant, which guarantees traditional relational
algebra operators must never observe a missing value in any attribute that they operate on directly (\Cref{sec:placement}). 
The decision to place imputation operators is driven by our cost model (\Cref{sec:cost-model}), which characterizes a plan based on estimates of the 
quality of the imputations performed and the runtime performance. Our planning algorithm is
agnostic to the type of imputation used (\Cref{sec:imputation}).
Finally, we show that while our algorithm is exponential in the number of joins (\Cref{sec:complexity}), planning times are not a concern in practice.

>>>>

\section{Algorithm}
\label{sec:algorithm}
In order to correctly plan around missing values, we first extend the set of relational algebra operators (selection $\sigma$, projection $\pi$, join $\bowtie$, and group-by/aggregate $g$) 
with two new operators (\Cref{sec:operators}). We then define the search space (\Cref{sec:search-space}) of plans to encompass non-trivial operations, such as joins and 
aggregations. Soundness of query execution is provided by our main design invariant, which guarantees traditional relational
algebra operators must never observe a missing value in any attribute that they operate on directly (\Cref{sec:placement}). 
The decision to place imputation operators is driven by our cost model (\Cref{sec:cost-model}), which characterizes a plan based on estimates of the 
quality of the imputations performed and the runtime performance. Our planning algorithm is
agnostic to the type of imputation used (\Cref{sec:imputation}).
Finally, we show that while our algorithm is exponential in the number of joins (\Cref{sec:complexity}), planning times are not a concern in practice.

\begin{table}
  \centering
  \caption{Notation used in Section~\ref{sec:algorithm}.}
  \begin{tabular}{ll}
    \toprule
    Operator & Description \\
    \midrule
    $\textsc{Attr}(\cdot)$ & Attributes used in a predicate. \\
    $\textsc{Time}(\cdot)$ & Estimated time to run a plan. \\
    $\textsc{Penalty}(\cdot)$ & Quality penalty of a query plan. \\
    $\textsc{Card}(\cdot)$ & Estimated cardinality of a plan. \\
    $\widehat{Var}(\cdot)$ & Estimated variance of a column. \\
    $\textsc{Dirty}(\cdot)/\textsc{Clean}(\cdot)$ & Dirty/clean attributes in a plan. \\
    $Q[\cdot]$ & Plan cache access operator. \\
    $\cdot \triangleleft \cdot$ & Plan cache insertion operator. \\
    \bottomrule
  \end{tabular}
\end{table}
